A subjective Bayesian framework for synthesizing deep uncertainties in climate risk management

### 2022-06-22 papers/earths-future/submission.tex
Format to prepare for Earth's Future submission

\documentclass[11pt]{article}

@@

% fonts and appearance
\usepackage{amsmath, amsfonts, physics, siunitx, nicefrac}
\usepackage[american]{babel} 
\usepackage[T1]{fontenc} % improved font encoding
\usepackage[ttscale=0.8]{libertine}
\usepackage{fontawesome5}
\usepackage[format=plain, textfont=it]{caption}

% page size and margins
\usepackage{geometry}
\geometry{letterpaper,top=1in, bottom=1in, left=1in, right=2in}

% footer
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage[en-US]{datetime2}
\fancyhf{}
\fancyhead[L]{CONFIDENTIAL DRAFT by J. Doss-Gollin \& K. Keller}
\fancyhead[R]{\DTMnow}
\fancyfoot[R]{page~\thepage~of~\pageref{LastPage}}
\pagestyle{fancy}
\setlength{\headheight}{15pt}

% TO DO NOTES
\usepackage{xcolor} % list of colors at https://en.wikibooks.org/wiki/LaTeX/Colors
\definecolor{giallo}{HTML}{F0BC42} % https://teamcolorcodes.com/a-s-roma-color-codes/
\definecolor{rosso}{HTML}{8E1F2F}
\definecolor{grigio}{HTML}{CACACC}
\definecolor{nero}{HTML}{000000}
\usepackage[textsize=scriptsize]{todonotes}
\setlength{\marginparwidth}{1.5in}
\newcommand{\james}[1]{\todo[color=giallo, textcolor=nero]{\textbf{ATTN James:~}#1}} % if desired create a custom command for each author
\newcommand{\klaus}[1]{\todo[color=rosso, textcolor=grigio]{\textbf{ATTN Klaus:~}#1}}

@@

% better lists
\usepackage[inline]{enumitem}
\setlist{nosep}

% authors
\usepackage{authblk}
\title{A subjective Bayesian framework for synthesizing deep uncertainties in climate risk management}
\author[1]{James Doss-Gollin}
\author[2]{Klaus Keller}
\affil[1]{Department of Civil and Environmental Engineering, Rice University}
\affil[2]{Thayer School of Engineering, Dartmouth College}
\renewcommand*{\Affilfont}{\normalsize\normalfont}

% ACRONYMS
\usepackage[acronym, nopostdot, nonumberlist, shortcuts, numberedsection, nogroupskip,]{glossaries}
\newacronym{bfe}{BFE}{base flood elevation}
\newacronym{cmip}{CMIP}{the Coupled Model Intercomparison Project}
\newacronym{dmdu}{DMDU}{decision making under deep uncertainty}
\newacronym{fema}{FEMA}{the Federal Emergency Management Agency}
\newacronym{gev}{GEV}{generalized extreme value}
\newacronym{hazus}{HAZUS}{Hazard U.S.}
\newacronym{iid}{IID}{independent and identically distributed}
\newacronym{ipcc}{IPCC}{International Panel on Climate Change}
\newacronym{msl}{MSL}{mean relative sea level}
\newacronym{noaa}{NOAA}{the National Oceanic and Atmospheric Administration}
\newacronym{pdf}{PDF}{probability density function}
\newacronym{rcp}{RCP}{representative concentration pathway}
\newacronym{rdm}{RDM}{robust decision making}
\newacronym{slr}{SLR}{sea level rise}
\newacronym{ssp}{SSP}{shared socio-economic pathway}
\newacronym[]{usace}{USACE}{United States Army Corps of Engineers}
\newacronym[]{usgs}{USGS}{United States Geological Survey}
\newacronym[plural=SOWs,firstplural=states of the world (SOWs),descriptionplural=states of the world]{sow}{SOW}{state of the world}
\makeglossaries % required to make the list of acronyms

\usepackage{xspace}
\makeatletter
\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
\def\eg{\emph{e.g}\onedot} \def\Eg{\emph{E.g}\onedot}
\def\ie{\emph{i.e}\onedot} \def\Ie{\emph{I.e}\onedot}
\def\etc{\emph{etc}\onedot} \def\vs{\emph{vs}\onedot}

\usepackage{xspace}
\makeatletter
\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
\newcommand{\usd}[1]{\SI{#1}[\$]{}}
\def\eg{\emph{e.g}\onedot} \def\Eg{\emph{E.g}\onedot}
\def\ie{\emph{i.e}\onedot} \def\Ie{\emph{I.e}\onedot}
\def\etc{\emph{etc}\onedot} \def\vs{\emph{vs}\onedot}

% use biblatex
\usepackage{csquotes}
\usepackage[
  backend=biber,
  doi=true,
  url=false,
  isbn=false,
  style=authoryear-comp,
  natbib=true,
  backref=false,
  maxbibnames=10,
  maxcitenames=2,
  uniquename=false,
  uniquelist=false,
  sorting=nyt,
  giveninits=true,
]{biblatex}
\renewbibmacro{in:}{}
\AtEveryBibitem{\clearfield{month}\clearfield{day}\clearfield{pages}\clearlist{language}}
\addbibresource{library.bib}

% load this last
\usepackage[hidelinks]{hyperref}
\usepackage{cleveref}

\begin{document}
\maketitle
\thispagestyle{empty}

\begin{abstract}
    Current practice in engineering, infrastructure design, and local governance relies heavily upon standards that specify particular design events or conditions that buildings and infrastructure should safely withstand.
    The
\end{abstract}

Key points
\begin{enumerate}
    \item
    \item
    \item
\end{enumerate}

\clearpage
\section{Introduction}\label{sec:introduction}

Many critical infrastructure services around the world are aging and inadequate for a changing climate \citep[\eg,][]{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020}.
Looking to the future, it is expected that changes to regulation, economics and finance, patterns of population and infrastructure use, as well as climate, will further stress infrastructure systems.
This motivates the question: which possible futures should infrastructure systems and components be designed for?
The answer to this question depends in part on values: intrinsic trade-offs between safety, performance, cost, and other objectives depend on context and stakeholder preferences \citep{keller_management:2021} and are often regulated by statue or industry guidelines.
For example, hospitals and critical infrastructure are generally designed to a higher standard of risk protection than ordinary buildings.
At the same time, performance depends upon future conditions, and so decisions about which scenario to design for are necessarily subject to implicit or explicit assumptions about the likelihood or possibility of different future conditions.

@@

Current practice in engineering, infrastructure design, and local governance relies heavily upon standards that specify particular design events or conditions that buildings and infrastructure should safely withstand \citep{asce_7-10:2013,bruneau_multihazard:2017}.
These are often, though not always, informed by probabilistic analysis of relevant data.
For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.
These trigger specific floodplain regulations, such as a requirement for homes in the 100-year floodplain with federally backed mortgages to be covered by flood insurance \citep{kousky_voucher:2014}.
Additionally, local building codes \citep[based on guidance such as][]{asce_24-05:2006,FEMA_p-55:2011} may require new construction in flood zones to be elevated freeboard above a nominal \gls{bfe}.
Design standards are also used to design large-scale infrastructure.
For example, some levees in the Netherlands are required to be designed such that the annual probability of failure is less than 1 in \num{4000} \citep{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project for  Galveston Bay was designed by setting the nominal annual probability of overtopping to 1\% \citep[Appendix D.,~p.~2-59]{USACE_coastal:2021}.

There are many advantages to standards-based design.
In particular, these heuristics are scalable and explainable, do not require complex analysis to apply, and are fair in at least a procedural sense.
However, there are also limitations.
In particular, this sort of one-size-fits-all guidance may not be an efficient or desirable way to balance tradeoffs.
This has motivated economically informed approaches, like risk-based design and cost benefit analysis \citep{eijgenraam_flooding:2014,vandantzig_dike:1956,xian_elevation:2017}, that place ``a strong emphasis upon a proportionate response to risk, so that the amount invested in risk reduction is in proportion to the magnitude of the risk and the cost-effectiveness with which that risk may be reduced'' \citep{merz_fluvial:2010}.
These quantitative cost-benefit analyses also rely on probabilistic descriptions of relevant hazard, and may be particularly sensitive to representation of tail probabilities.

Currently used methods for estimating these \glspl{pdf} emphasize data-driven, nominally objective methods that can be applied consistently across locations.
For example, \gls{usgs} Bulletin 17C specifies procedures for estimating flood frequency and \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall.
Among several statistical assumptions these analyses make that have been recently called into question is stationarity \citep[the assumption that past and future hazard come from the same \gls{pdf}; see][for a review]{merz_review:2014}.
For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \citep{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards are consistent with observations \citep[see][for a comprehensive summary]{ipcc_impacts:2022}.\klaus{I'm citing the technical summary which is 96 pages -- there are about 100 authors listed. Any suggestion for how to cite best?}
While some methods have been proposed for incorporating trends into these analyses \citep[see][for a review]{Salas:2018ge}, these assume specific forms of nonstationarity which may not fully represent physical processes or true levels of uncertainty \citep{DossGollin:2019,Montanari:2014hl,Serinaldi:2015bq}.
Ultimately, the difficulty of incorporating nonstationarity into existing frameworks has led to continued reliance on stationarity despite recognition of the associated limitations (see \cite{bulletin17c:2019}, p.~2, or \cite{atlas14_texas:2018} p.~A.4-42).

@@

Projecting nonstationary hazard is difficult because many future hazards depend on intrinsically unpredictable human decisions (\eg, the rate of future greenhouse gas emissions) or on physical processes that are poorly constrained by existing data \citep[\eg, collapse of the West Antarctic ice sheet; see][]{deconto_antarctica:2016}.
In other words, they are deeply uncertain \citep{keller_management:2021,walker_deep:2013,lempert_complex:2002}.
These deeply uncertain nonstationary hazards challenge not only the existing stationary estimates but, more fundamentally, the premise that objective estimates of future hazard exist and can be estimated empirically.

Recognizing the challenges of \gls{dmdu}, many frameworks for identifying robust decisions have been proposed.
Most emphasize the use of models in an exploratory (``what-if'') framework to learn about interactions between decisions and system dynamics \citep{bankes:1993}.
For example, robust decision making \citep{lempert_shaping:2003} evaluates models over large ensembles of possible futures to assess the performance of different policies under each, then applies statistical analysis to identify the conditions under which particular policies perform well or poorly.
When the decision space is complex, many analyses use policy search (often using multiobjective optimization tools) to identify promising actions \citep{kasprzyk:2013,kasprzyk_denovo:2012,hadka_mordm:2015}.
In addition, many studies formally quantify robustness to deep uncertainties \citep{herman:2015,mcphail_robustness:2019} and use this as a criterion for policy comparison.

Although these methods for \gls{dmdu} have proven valuable in a wide range of settings, they still require necessarily subjective assessments about the likelihood of future conditions.
Even exploratory models require the analyst to choose which uncertainties are considered and how they are sampled.
When analyses use metrics that integrate performance over many possible futures, whether to compute sensitivities, estimate expectations, or compute robustness metrics, they necessarily make assumptions about the likelihood of different futures.
For example, uniform priors are often justified by deferring to Laplace's principle of insufficient reason \citep[see][p.~135]{stigler_uncertainty:1986}, but even this choice makes strong assumptions that are sensitive to factors such as parameterization \citep[p.54]{gelman_bda3:2014}.
While one interpretation of deep uncertainty is that that specifying a joint \gls{pdf} over inputs is inappropriate, assumptions about the ranges and independence of parameters to sample are just as subjective as the choice of probability distribution \citep{schneider_scenarios:2002,quinn_exploratory:2020} and, indeed, can be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
This motivates the development of decision analytic frameworks that draw from the strengths of \gls{dmdu} methods such as exploratory modeling, vulnerability assessment, robustness checks, and iterative stakeholder critique, but that embrace the reality that assumptions about the future are inescapable.

@@

We are motivated by parallels between decision making under deep uncertainty and the statistical problem of model selection.
\citet{oreskes_verification:1994} argues that because natural systems are never closed and model results are never unique, validation and verification of models representing these systems is necessarily qualitative and subjective.
In the literature on decision making under deep uncertainty, this has led to recognition of the need to develop strategies that are robust to model imperfections.
We find parallels to this challenge in the Bayesian literature on model selection, particularly in the ``$\mathcal{M}$-open'' case in which there is no ``true'' model to identify and all models are imperfect representations of reality.
This is a theoretical contrast to the ``$\mathcal{M}$''-closed case in which a true model can be identified and written down but is one amongst finitely many models from which an analyst has to choose as well as the ``$\mathcal{M}$''-complete case in which a true model exists but is inaccessible in the sense that even though it can be conceptualized it cannot be written down \citep{bernardo_bayesian:1994}.
Whereas model selection in the $\mathcal{M}$-closed case seeks to asymptotically identify the true model from a set of candidate models, model selection in the $\mathcal{M}$-open case emphasizes a subjective view, both of probability itself and of the modeling process, for which probability offers a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \cite{gelman_philosophy:2013}, \cite{ramsey_probability:2016}, or \cite{jaynes_probability:2003} for a discussion of Bayesian philosophy and \cite{Piironen:2017eh} for a technical discussion of methods for model selection).
From a practical perspective, model selection in the $\mathcal{M}$-open case emphasizes building models iteratively, simulating the consequences of these models, and subsequently using these simulations to critique and improve them \citep{gelman_workflow:2020}.

In this paper we offer a first conceptual step towards bridging the fields of \gls{dmdu} and Bayesian model building.
We consider a didactic case study of whether to elevate a hypothetical house, and if so how high, as a specific example of a decision problem subject to both shallow (storm surge) and deep (\gls{slr}) uncertainties.
Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,rozer_coping:2016,mobley_mitigation:2020,aerts_cost:2018}, and both local building codes \citep{asce_7-10:2013,bruneau_multihazard:2017,asce_24-05:2006} and federal policy \citep{FEMA_p-55:2011} require elevation in some cases.
Guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent research has demonstrated that neglecting uncertainty in the cost-benefit analysis can lead to poor decisions \citep{zarekarizi_suboptimal:2020}.
Focusing on deep uncertainty in \gls{slr} over the 70\james{Need to re-run codes with 70 rather than 71} year design life of a hypothetical house, we seek to answer the research question ``\emph{how can deep uncertainties be transparently synthesized for decision analysis?}''

@@

A first analytical step is to use the model in an ``exploratory'' mode.
Exploratory modeling is averse to making explicit assumptions about the likelihood of different \glspl{sow} and instead seek to generate new knowledge \citep{bankes:1993} by systematically exploring a large number of possible futures, emphasizing interactions between different uncertainties \citep{reed_msdbook:2022}.
Exploratory modeling is often paired with analyses that identify relevant scenarios  \citep{lamontagne_discovery:2018,groves_scenarios:2007} or summarize a system's response to forcing \citep{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
Despite the aversion to strong assumptions about the likelihood of different futures, subjective modeling decisions such as the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} strongly influence results \citep{quinn_exploratory:2020}.

>>>>

\documentclass[draft]{agujournal2019}

%% required packages
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}

@@

% fonts and appearance
\usepackage{amsmath, amsfonts, physics, siunitx, nicefrac}

@@

% ACRONYMS
\usepackage[acronym, nopostdot, nonumberlist, shortcuts, numberedsection, nogroupskip,]{glossaries}
\newacronym{bfe}{BFE}{base flood elevation}
\newacronym{cmip}{CMIP}{the Coupled Model Intercomparison Project}
\newacronym{dmdu}{DMDU}{decision making under deep uncertainty}
\newacronym{fema}{FEMA}{the Federal Emergency Management Agency}
\newacronym{gev}{GEV}{generalized extreme value}
\newacronym{hazus}{HAZUS}{Hazard U.S.}
\newacronym{iid}{IID}{independent and identically distributed}
\newacronym{ipcc}{IPCC}{International Panel on Climate Change}
\newacronym{msl}{MSL}{mean relative sea level}
\newacronym{noaa}{NOAA}{the National Oceanic and Atmospheric Administration}
\newacronym{pdf}{PDF}{probability density function}
\newacronym{rcp}{RCP}{representative concentration pathway}
\newacronym{rdm}{RDM}{robust decision making}
\newacronym{slr}{SLR}{sea level rise}
\newacronym{ssp}{SSP}{shared socio-economic pathway}
\newacronym[]{usace}{USACE}{United States Army Corps of Engineers}
\newacronym[]{usgs}{USGS}{United States Geological Survey}
\newacronym[plural=SOWs,firstplural=states of the world (SOWs),descriptionplural=states of the world]{sow}{SOW}{state of the world}

\usepackage{xspace}
\makeatletter
\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
\def\eg{\emph{e.g}\onedot} \def\Eg{\emph{E.g}\onedot}
\def\ie{\emph{i.e}\onedot} \def\Ie{\emph{I.e}\onedot}
\def\etc{\emph{etc}\onedot} \def\vs{\emph{vs}\onedot}
\newcommand{\usd}[1]{\SI{#1}[\$]{}}

% cite things in supplemental info file
\usepackage{xr}
\externaldocument{supplemental}

% this goes last
\usepackage{cleveref}

\linenumbers
%%%%%%%
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%

\draftfalse

%% Enter journal name below.
%% Choose from this list of Journals:
%
% JGR: Atmospheres
% JGR: Biogeosciences
% JGR: Earth Surface
% JGR: Oceans
% JGR: Planets
% JGR: Solid Earth
% JGR: Space Physics
% Global Biogeochemical Cycles
% Geophysical Research Letters
% Paleoceanography and Paleoclimatology
% Radio Science
% Reviews of Geophysics
% Tectonics
% Space Weather
% Water Resources Research
% Geochemistry, Geophysics, Geosystems
% Journal of Advances in Modeling Earth Systems (JAMES)
% Earth's Future
% Earth and Space Science
% Geohealth
%
% ie, \journalname{Water Resources Research}

\journalname{Earth's Future}

\begin{document}

%% ------------------------------------------------------------------------ %%
%  Title
%
% (A title should be specific, informative, and brief. Use
% abbreviations only if they are defined in the abstract. Titles that
% start with general keywords then specific terms are optimized in
% searches)
%
%% ------------------------------------------------------------------------ %%

% Example: \title{This is a test title}

\title{A subjective Bayesian framework for synthesizing deep uncertainties in climate risk management}

%% ------------------------------------------------------------------------ %%
%
%  AUTHORS AND AFFILIATIONS
%
%% ------------------------------------------------------------------------ %%

% Authors are individuals who have significantly contributed to the
% research and preparation of the article. Group authors are allowed, if
% each author in the group is separately identified in an appendix.)

% List authors by first name or initial followed by last name and
% separated by commas. Use \affil{} to number affiliations, and
% \thanks{} for author notes.
% Additional author notes should be indicated with \thanks{} (for
% example, for current addresses).

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}

\authors{James Doss-Gollin\affil{1}, Klaus Keller\affil{2}}
\affiliation{1}{Department of Civil and Environmental Engineering, Rice University}
\affiliation{2}{Thayer School of Engineering, Dartmouth College}

%% Corresponding Author:
% Corresponding author mailing address and e-mail address:

% (include name and email addresses of the corresponding author.  More
% than one corresponding author is allowed in this LaTeX file and for
% publication; but only one corresponding author is allowed in our
% editorial system.)

% Example: \correspondingauthor{First and Last Name}{email@address.edu}

\correspondingauthor{James Doss-Gollin}{jdossgollin@rice.edu}

%% Keypoints, final entry on title page.

%  List up to three key points (at least one is required)
%  Key Points summarize the main points and conclusions of the article
%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences

% Example:
% \begin{keypoints}
% \item	List up to three key points (at least one is required)
% \item	Key Points summarize the main points and conclusions of the article
% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
% \end{keypoints}

\begin{keypoints}
  \item Decisions about how to manage climate risks are often made under deep uncertainty
  \item We introduce a Bayesian framework to transparently synthesize and characterize deep uncertainties with the goal to support decision-making
  \item We demonstrate the framework using a simple case study of coastal flood risk management
\end{keypoints}

%% ------------------------------------------------------------------------ %%
%
%  ABSTRACT and PLAIN LANGUAGE SUMMARY
%
% A good Abstract will begin with a short description of the problem
% being addressed, briefly describe the new data or analyses, then
% briefly states the main conclusion(s) and how they are supported and
% uncertainties.

% The Plain Language Summary should be written for a broad audience,
% including journalists and the science-interested public, that will not have 
% a background in your field.
%
% A Plain Language Summary is required in GRL, JGR: Planets, JGR: Biogeosciences,
% JGR: Oceans, G-Cubed, Reviews of Geophysics, and JAMES.
% see http://sharingscience.agu.org/creating-plain-language-summary/)
%
%% ------------------------------------------------------------------------ %%

%% \begin{abstract} starts the second page

\begin{abstract}
  Projections of future climate risks can vary considerably between sources without a  consensus about their  relative probability.
  This deep uncertainty poses considerable communication and decision-analytical challenges.
  One such challenge is how to present trade-offs under deep uncertainty in a salient and interpretable manner.
  Some common approaches include analyzing a small subset of projections or invoking Laplace's principle of insufficient reason to justify a simple average.
  These approaches (i) can underestimate risks, (ii) provide little insights into which assumptions drive decision-relevant outcomes, and (iii) can nudge how people choose their decisions-making frameworks.
  Here we introduce and demonstrate a transparent Bayesian framework for synthesizing deep uncertainties to inform decision support.
  The first step of this workflow is to create an ensemble of simulations representing possible futures and analyze them through standard exploratory modeling techniques.
  Next, a small set of probability distributions representing subjective beliefs over possible futures is used to weight the scenarios.
  Finally, these weights are used to compute and characterize trade-offs, conduct robustness checks, and reveal implicit assumptions.
  We demonstrate the framework through a didactic case study analyzing how high to elevate a house to manage coastal flood risks.
\end{abstract}

\section*{Plain Language Summary}

What are sound strategies to manage risks driven by climatic changes?
Addressing this question is complicated by the large uncertainties surrounding projections of the coupled natural-human systems.
These uncertainties often arise from choices experts have to make, for example about technical detail on how to model future water levels.
Different experts can disagree about these choices and provide different plausible projections.
Analyzing decisions in such a situation of deep uncertainty poses nontrivial challenges.
For one, just picking one of these projections or just an average can lead to poor decisions.
On the other extreme, communicating results separately for all projections can overwhelm decision-makers.
To make matters worse, typical approaches to this problem are mostly silent on what assumptions make a difference for the decisions at hand.
We develop and demonstrate a framework to address these challenges.
The framework provides a transparent approach (i) to combine a large number of deeply uncertain projections to a more understandable and smaller sample and (ii) to provide insights as to how assumptions and modeling choices made by experts influence decisions.
We demonstrate the approach with a relatively simple example question of how high to elevate a house in the face of deeply uncertain projections about future water levels.

%% ------------------------------------------------------------------------ %%
%
%  TEXT
%
%% ------------------------------------------------------------------------ %%

\section{Introduction}\label{sec:introduction}

Many critical infrastructure services around the world are aging and inadequate for a changing climate \cite<\eg,>[]{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020}.
Looking to the future, it is expected that changes to regulation, economics and finance, patterns of population and infrastructure use, as well as climate, will further stress infrastructure systems.
This motivates the question: which possible futures should infrastructure systems and components be designed for?
The answer to this question depends in part on values: intrinsic trade-offs between safety, performance, cost, and other objectives depend on context and stakeholder preferences \cite{keller_management:2021} and are often regulated by statue or industry guidelines.
For example, hospitals and critical infrastructure are generally designed to a higher standard of risk protection than ordinary buildings.
At the same time, performance depends upon future conditions, and so decisions about which scenario to design for are necessarily subject to implicit or explicit assumptions about the likelihood or possibility of different future conditions.

@@

Current practice in engineering, infrastructure design, and local governance relies heavily upon standards that specify particular design events or conditions that buildings and infrastructure should safely withstand \cite{asce_7-10:2013,bruneau_multihazard:2017}.
These are often, though not always, informed by probabilistic analysis of relevant data.
For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.
These trigger specific floodplain regulations, such as a requirement for homes in the 100-year floodplain with federally backed mortgages to be covered by flood insurance \cite{kousky_voucher:2014}.
Additionally, local building codes \cite<based on guidance such as>[]{asce_24-05:2006,FEMA_p-55:2011} may require new construction in flood zones to be elevated freeboard above a nominal \gls{bfe}.
Design standards are also used to design large-scale infrastructure.
For example, some levees in the Netherlands are required to be designed such that the annual probability of failure is less than 1 in \num{4000} \cite{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project for  Galveston Bay was designed by setting the nominal annual probability of overtopping to 1\% \cite[Appendix D.,~p.~2-59]{USACE_coastal:2021}.

There are many advantages to standards-based design.
In particular, these heuristics are scalable and explainable, do not require complex analysis to apply, and are fair in at least a procedural sense.
However, there are also limitations.
In particular, this sort of one-size-fits-all guidance may not be an efficient or desirable way to balance tradeoffs.
This has motivated economically informed approaches, like risk-based design and cost benefit analysis \cite{eijgenraam_flooding:2014,vandantzig_dike:1956,xian_elevation:2017}, that place ``a strong emphasis upon a proportionate response to risk, so that the amount invested in risk reduction is in proportion to the magnitude of the risk and the cost-effectiveness with which that risk may be reduced'' \cite{merz_fluvial:2010}.
These quantitative cost-benefit analyses also rely on probabilistic descriptions of relevant hazard, and may be particularly sensitive to representation of tail probabilities.

Currently used methods for estimating these \glspl{pdf} emphasize data-driven, nominally objective methods that can be applied consistently across locations.
For example, \gls{usgs} Bulletin 17C specifies procedures for estimating flood frequency and \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall.
Among several statistical assumptions these analyses make that have been recently called into question is stationarity \cite<the assumption that past and future hazard come from the same \gls{pdf}; see>[for a review]{merz_review:2014}.
For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \cite{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards are consistent with observations \cite<see>[for a comprehensive summary]{ipcc_impacts:2022}.
While some methods have been proposed for incorporating trends into these analyses \cite<see>[for a review]{Salas:2018ge}, these assume specific forms of nonstationarity which may not fully represent physical processes or true levels of uncertainty \cite{DossGollin:2019,Montanari:2014hl,Serinaldi:2015bq}.
Ultimately, the difficulty of incorporating nonstationarity into existing frameworks has led to continued reliance on stationarity despite recognition of the associated limitations (see \cite{bulletin17c:2019}, p.~2, or \cite{atlas14_texas:2018} p.~A.4-42).

@@

Projecting nonstationary hazard is difficult because many future hazards depend on intrinsically unpredictable human decisions (\eg, the rate of future greenhouse gas emissions) or on physical processes that are poorly constrained by existing data \cite<\eg, collapse of the West Antarctic ice sheet; see>[]{deconto_antarctica:2016}.
In other words, they are deeply uncertain \cite{keller_management:2021,walker_deep:2013,lempert_complex:2002}.
These deeply uncertain nonstationary hazards challenge not only the existing stationary estimates but, more fundamentally, the premise that objective estimates of future hazard exist and can be estimated empirically.

Recognizing the challenges of \gls{dmdu}, many frameworks for identifying robust decisions have been proposed.
Most emphasize the use of models in an exploratory (``what-if'') framework to learn about interactions between decisions and system dynamics \cite{bankes:1993}.
For example, robust decision making \cite{lempert_shaping:2003} evaluates models over large ensembles of possible futures to assess the performance of different policies under each, then applies statistical analysis to identify the conditions under which particular policies perform well or poorly.
When the decision space is complex, many analyses use policy search (often using multiobjective optimization tools) to identify promising actions \cite{kasprzyk:2013,kasprzyk_denovo:2012,hadka_mordm:2015}.
In addition, many studies formally quantify robustness to deep uncertainties \cite{herman:2015,mcphail_robustness:2019} and use this as a criterion for policy comparison.

Although these methods for \gls{dmdu} have proven valuable in a wide range of settings, they still require necessarily subjective assessments about the likelihood of future conditions.
Even exploratory models require the analyst to choose which uncertainties are considered and how they are sampled.
When analyses use metrics that integrate performance over many possible futures, whether to compute sensitivities, estimate expectations, or compute robustness metrics, they necessarily make assumptions about the likelihood of different futures.
For example, uniform priors are often justified by deferring to Laplace's principle of insufficient reason \cite<see>[p.~135]{stigler_uncertainty:1986}, but even this choice makes strong assumptions that are sensitive to factors such as parameterization \cite[p.54]{gelman_bda3:2014}.
While one interpretation of deep uncertainty is that that specifying a joint \gls{pdf} over inputs is inappropriate, assumptions about the ranges and independence of parameters to sample are just as subjective as the choice of probability distribution \cite{schneider_scenarios:2002,quinn_exploratory:2020} and, indeed, can be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
This motivates the development of decision analytic frameworks that draw from the strengths of \gls{dmdu} methods such as exploratory modeling, vulnerability assessment, robustness checks, and iterative stakeholder critique, but that embrace the reality that assumptions about the future are inescapable.

@@

We are motivated by parallels between decision making under deep uncertainty and the statistical problem of model selection.
\citeA{oreskes_verification:1994} argues that because natural systems are never closed and model results are never unique, validation and verification of models representing these systems is necessarily qualitative and subjective.
In the literature on decision making under deep uncertainty, this has led to recognition of the need to develop strategies that are robust to model imperfections.
We find parallels to this challenge in the Bayesian literature on model selection, particularly in the ``$\mathcal{M}$-open'' case in which there is no ``true'' model to identify and all models are imperfect representations of reality.
This is a theoretical contrast to the ``$\mathcal{M}$''-closed case in which a true model can be identified and written down but is one amongst finitely many models from which an analyst has to choose as well as the ``$\mathcal{M}$''-complete case in which a true model exists but is inaccessible in the sense that even though it can be conceptualized it cannot be written down \cite{bernardo_bayesian:1994}.
Whereas model selection in the $\mathcal{M}$-closed case seeks to asymptotically identify the true model from a set of candidate models, model selection in the $\mathcal{M}$-open case emphasizes a subjective view, both of probability itself and of the modeling process, for which probability offers a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \cite{gelman_philosophy:2013}, \cite{ramsey_probability:2016}, or \cite{jaynes_probability:2003} for a discussion of Bayesian philosophy and \cite{Piironen:2017eh} for a technical discussion of methods for model selection).
From a practical perspective, model selection in the $\mathcal{M}$-open case emphasizes building models iteratively, simulating the consequences of these models, and subsequently using these simulations to critique and improve them \cite{gelman_workflow:2020}.

In this paper we offer a first conceptual step towards bridging the fields of \gls{dmdu} and Bayesian model building.
We consider a didactic case study of whether to elevate a hypothetical house, and if so how high, as a specific example of a decision problem subject to both shallow (storm surge) and deep (\gls{slr}) uncertainties.
Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \cite{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,rozer_coping:2016,mobley_mitigation:2020,aerts_cost:2018}, and both local building codes \cite{asce_7-10:2013,bruneau_multihazard:2017,asce_24-05:2006} and federal policy \cite{FEMA_p-55:2011} require elevation in some cases.
Guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \cite{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent research has demonstrated that neglecting uncertainty in the cost-benefit analysis can lead to poor decisions \cite{zarekarizi_suboptimal:2020}.
Focusing on deep uncertainty in \gls{slr} over the 70 year design life of a hypothetical house, we seek to answer the research question ``\emph{how can deep uncertainties be transparently synthesized for decision analysis?}''

@@

A first analytical step is to use the model in an ``exploratory'' mode.
Exploratory modeling is averse to making explicit assumptions about the likelihood of different \glspl{sow} and instead seek to generate new knowledge \cite{bankes:1993} by systematically exploring a large number of possible futures, emphasizing interactions between different uncertainties \cite{reed_msdbook:2022}.
Exploratory modeling is often paired with analyses that identify relevant scenarios  \cite{lamontagne_discovery:2018,groves_scenarios:2007} or summarize a system's response to forcing \cite{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
Despite the aversion to strong assumptions about the likelihood of different futures, subjective modeling decisions such as the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} strongly influence results \cite{quinn_exploratory:2020}.
